---
title: "Model Mis-specification Monte Carlo Simulation"
author: "Jianing Li"
date: "2026-01-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Overview

This document presents a Monte Carlo simulation study examining
the finite-sample bias, variance, and MSE of OLS estimators under
model misspecification. The simulation focuses on omitted-variable
bias under alternative correlation structures and sample sizes.
All simulations are implemented in R and are fully reproducible.

Note: This simulation was originally developed as part of a graduate
econometrics course and has been lightly reorganized for clarity.

# Model mis-specification Mote Carlo simulations

True Model: $y_t=\alpha+\beta x_{1t}+\gamma x_{2t}+u_t$

Under-specified Model: $y_t=\alpha+\beta x_{1t}+e_t$ for $t=1,2,...,621$.

## (1)  Generate the data under the true model

```{r}
library(MASS) # For multivariate models
library(ggplot2)
library(reshape2)

# Set seed so Professor Xu can reproduce the same thing :D
set.seed(622)

n = 621
mu = matrix(c(0, 1), nrow = 2, byrow = TRUE)

sigma1 = 1
sigma2 = 4
alpha = 0.6
beta = 0.2
gamma = 0.1
rho = 0.126

Sigma = matrix(c(sigma1^2, rho*sigma1*sigma2, rho*sigma1*sigma2, sigma2^2), nrow = 2, byrow = TRUE)

u = rnorm(n, 0, 1)

X = mvrnorm(n, mu = c(0,1), Sigma = Sigma)
x1 = X[, 1]
x2 = X[, 2]

y = alpha + beta*x1 + gamma*x2 + u
```

## (2) Run regressions

1. True model

```{r}
true_model = lm(y ~ x1 + x2)
true_model_results = coef(true_model)
true_model_results
```

2. Under-specified model

```{r}
under_model = lm(y ~ x1)
under_model_results = coef(under_model)
under_model_results
```

## (3) Repeat 621 times and Calculate the mean, bias, variance, minimum, maximum and mean square error

```{r}
stats_calculation = function(est, value){
  bias = colMeans(est) - value
  variance = apply(est, 2, var)
  mse = bias ^ 2 + variance
  
results = data.frame(
  Parameter = colnames(est),
  True_Value = value,
  Mean = colMeans(est),
  Bias = bias,
  Variance = variance,
  MSE = mse,
  Min = apply(est, 2, min),
  Max = apply(est, 2, max)
)

return(results)
}

simulation = function(rho_value, n_obs, n_sims) {
  n = n_obs
  mu = matrix(c(0, 1), nrow = 2, byrow = TRUE)
  sigma1 = 1
  sigma2 = 4
  alpha = 0.6
  beta = 0.2
  gamma = 0.1
  
  Sigma = matrix(c(sigma1^2, rho*sigma1*sigma2,
                   rho_value*sigma1*sigma2, sigma2^2), nrow = 2,
                 byrow = TRUE)

  true_model_results = matrix(NA, nrow = n_sims, ncol = 3)
  colnames(true_model_results) <- c("alpha_hat", "beta_hat", "gamma_hat")
  
  under_model_results = matrix(NA, nrow = n_sims, ncol = 2)
  colnames(under_model_results) = c("alpha_hat", "beta_hat")
  
  for (i in 1:n_sims) {
    set.seed(i)
    u = rnorm(n, 0, 1)

    X = mvrnorm(n, mu = c(0,1), Sigma = Sigma)
    x1 = X[, 1]
    x2 = X[, 2]
    y = alpha + beta*x1 + gamma*x2 + u
  
    true_model = lm(y ~ x1 + x2)
    true_model_results[i, ] = coef(true_model)

    under_model = lm(y ~ x1)
    under_model_results[i, ] = coef(under_model)
  }
  
   true_summary = stats_calculation(true_model_results, 
                                 c(alpha, beta,
                                   gamma))
  
  under_summary =   
    stats_calculation(under_model_results, 
                      c(alpha, beta))
  
  return(list(
    true_model = true_model_results,
    under_model = under_model_results,
    true_summary = true_summary,
    under_summary = under_summary,
    rho = rho_value
  ))
}

create_summary_table = function(results, model_name) {
  true_df = results$true_summary
  under_df = results$under_summary
  
  combined <- rbind(
    cbind(Model = "True", true_df),
    cbind(Model = "Under-specified", 
          rbind(under_df, 
                data.frame(Parameter = "gamma_hat", True_Value = 0.1, 
                           Mean = NA, Bias = NA, Variance = NA, MSE = NA, 
                           Min = NA, Max = NA)))
  )
  
  formatted = combined[, c("Model", "Parameter", "True_Value", "Mean", "Bias", "Variance", "MSE")]
  colnames(formatted) = c("Model", "Parameter", "True Value", "Mean", "Bias", "Variance", "MSE")
  
  formatted[, 3:7] = round(formatted[, 3:7], 4)
  
  cat(paste("\n", model_name, "\n"))
  cat("=====================\n")
  print(formatted, row.names = FALSE)
  cat("\n")
  
  return(formatted)
}

results_rho1 = simulation(rho_value = 0.126, n_obs = 621, n_sims = 621)

table_rho1 = create_summary_table(results_rho1, "SUMMARY TABLE: rho = 0.126")
table_rho1

create_comparison_plots = function(results, rho_value) {
  true_beta = data.frame(Value = results$true_model[, "beta_hat"], Model = "True Model")
  under_beta = data.frame(Value = results$under_model[, "beta_hat"], Model = "Under-specified")
  beta_data = rbind(true_beta, under_beta)
  
  true_alpha = data.frame(Value = results$true_model[, "alpha_hat"], Model = "True Model")
  under_alpha = data.frame(Value = results$under_model[, "alpha_hat"], Model = "Under-specified")
  alpha_data = rbind(true_alpha, under_alpha)
  
  p1 = ggplot(beta_data, aes(x = Value, fill = Model)) +
    geom_histogram(alpha = 0.7, position = "identity", bins = 30) +
    geom_vline(xintercept = 0.2, linetype = "dashed", color = "black", size = 1) +
    labs(title = paste("Distribution of Beta Estimates (rho =", rho_value, ")"),
         x = "Beta Estimate", y = "Frequency") +
    theme_minimal() +
    scale_fill_manual(values = c("True Model" = "blue", "Under-specified" = "red"))
  
  p2 = ggplot(alpha_data, aes(x = Value, fill = Model)) +
    geom_histogram(alpha = 0.7, position = "identity", bins = 30) +
    geom_vline(xintercept = 0.6, linetype = "dashed", color = "black", size = 1) +
    labs(title = paste("Distribution of Alpha Estimates (rho =", rho_value, ")"),
         x = "Alpha Estimate", y = "Frequency") +
    theme_minimal() +
    scale_fill_manual(values = c("True Model" = "blue", "Under-specified" = "red"))
  
  gamma_data = data.frame(Value = results$true_model[, "gamma_hat"])
  p3 <- ggplot(gamma_data, aes(x = Value)) +
    geom_histogram(fill = "green", alpha = 0.7, bins = 30) +
    geom_vline(xintercept = 0.1, linetype = "dashed", color = "black", size = 1) +
    labs(title = paste("Distribution of Gamma Estimates (rho =", rho_value, ")"),
         x = "Gamma Estimate", y = "Frequency") +
    theme_minimal()
  
  bias_data = data.frame(
    Model = c("True", "Under-specified"),
    Beta_Bias = c(results$true_summary$Bias[2], results$under_summary$Bias[2]),
    Alpha_Bias = c(results$true_summary$Bias[1], results$under_summary$Bias[1])
  )
  
  bias_melted = melt(bias_data, id.vars = "Model")
  p4 = ggplot(bias_melted, aes(x = Model, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
    labs(title = paste("Bias Comparison (rho =", rho_value, ")"),
         y = "Bias", fill = "Parameter") +
    theme_minimal() +
    scale_fill_manual(values = c("Beta_Bias" = "darkblue", "Alpha_Bias" = "darkred"))
  
  print(p1)
  print(p2)
  print(p3)
  print(p4)
  
  return(list(p1, p2, p3, p4))
}

plots_rho1 = create_comparison_plots(results_rho1, 0.126)
plots_rho1


```


## (4) Repeat 621 times with rho = 0.621

```{r}

results_rho2 = simulation(rho_value = 0.621, n_obs = 621, n_sims = 621)
table_rho2 = create_summary_table(results_rho2, "SUMMARY TABLE: rho = 0.621")
plots_rho2 = create_comparison_plots(results_rho2, 0.621)

```


## (5) Comments

Comparing to the true model, under-specified model shows a significant bias in $\beta$ estimates.  When $\rho = 0.126$ is lower, it produces small bias, when $\rho = 0.621$ is higher, it gives larger bias, which results in a poor performance.

From above, we assume that the cost of omitting variables increases with their correlation to included variables. High correlation makes misspecification worsen in bias and performance.
