---
title: "FNETS_RobustnessTest"
author: "Jianing Li"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# The codes is based on real_forecast_submission.R from paper FNETS: Factor-Adjusted Network Estimation and Forecasting for High-Dimensional Time Series by Mateeo Barigozzi, Haeran Cho, Dom Owens (2024), https://doi.org/10.1080/07350015.2023.2257270.

# The code will approximately need around 2 hours to run in total.  Specifically, FNETS with n=252 rolling window will need around 1 hr 30 mins and n=126 rolling window will need around 30-50 mins.

# Our modifications on the FNETS main body:

# 1. Removed the part of FARM methods which was the "baseline" method metioned in the paper for contrasting FNETS methods.

# 2. Removed fd forecast - full FNETS forecast via Dantzig selector.  It was used as robustness check in supplementary Appendix, we are going to do our robustness check via different rolling window (n=126) instead.

# 3. Removed fc forecast - Factor-based forecast because we are assessing the full FNETS forcasting performance instead of each factor's forecasting performance.

# 4. For time constraints, we only use restricted LASSO FNETS, because unrestricted one is also not the main target of the original paper.

# 5. For VAR(n), instead of original n=1:5, we only use n=1 because the original paper use VAR(1) to VAR(5) to do the sensitivity analysis, we don't need to do here.

## Start

library(glmnet)
library(fnets)
# load('blm.RData') ## Might need to manually input the file

```

```{r}

## Data Preparation

x0 <- t(old_clean[, -1]) # old_clean is the dataset name directly read from blm.RData with (46+1) * 3269 dimensions, -1 because the first column is the time dimension.

p <- dim(x0)[1] # the number of firms, 46 here.

dates <- as.Date(paste(old_clean[, 1]), format = '%Y%m%d')

n <- 252 # rolling window for 252 trading days, basically a year length of trading days.

kern.bw <- floor(4 * (n/log(n))^(1/3)) # In Section 3.2, Step 1: Factor Adjustment via Dynamic PCA, FNETS use kernel spectral density estimator and here is the bandwidth m in the function, it controls how many lags of the sample autocovariance are used when estimating the spectral density matrix.  Their theory gives an optimal choice of m should grow at the same rate as (n/log(n))^(1/3) as the sample size n increases.  The code ignores p because log(np) = log(n) when p is fixed, and times 4 for a scaling constant.

```

```{r}
## Regime split
tt_calm   <- which(old_clean$datadate >= 20040101 & old_clean$datadate <= 20061231)
tt_crisis <- which(old_clean$datadate >= 20070101 & old_clean$datadate <= 20091231)

tt_calm   <- tt_calm[tt_calm > n]
tt_crisis <- tt_crisis[tt_crisis > n]

cat('Calm:  ', length(tt_calm),   'days\n')
cat('Crisis:', length(tt_crisis), 'days\n')
```

# Assumptions Tests (Stationarity and Heavy-tailed)

```{r, warning=FALSE}
library(tseries)
library(moments)

# we set up a function for testing both stationarity and Heavy tails here

check_assumptions <- function(x, label) {
  p <- nrow(x)
  n <- ncol(x)
  cat("===============\n")
  cat("Period:", label, "\n")
  cat("Dimensions: p =", p, ", T =", n, "\n")
  cat("===============\n")

  # A1: Stationarity (ADF Test)
  adf_pvals <- apply(x, 1, function(z) adf.test(z)$p.value)
  cat("\n A1: Stationarity (ADF Test) \n")
  cat("% series stationary (p < 0.05):",
      round(mean(adf_pvals < 0.05) * 100, 1), "%\n")
  cat("Min p-value:", round(min(adf_pvals), 4), "\n")
  cat("Max p-value:", round(max(adf_pvals), 4), "\n")

  # A3: Heavy Tails (Sample kurtosis + Jarque-Bera + Anscombe-Glynn test)
  kurt     <- apply(x, 1, kurtosis) 
  jb_pvals <- apply(x, 1, function(z) jarque.bera.test(z)$p.value)
  ag_pvals <- apply(x, 1, function(z) anscombe.test(z)$p.value)
  
  cat("\n A3: Heavy Tails \n")
  cat("Mean kurtosis:", round(mean(kurt), 3), "\n")
  cat("% series JB reject normality (p<0.05):",
      round(mean(jb_pvals < 0.05) * 100, 1), "%\n")
  cat("% series AG reject kurtosis=3 (p<0.05):",
      round(mean(ag_pvals < 0.05) * 100, 1), "%\n")
}

check_assumptions(x0[, tt_calm],   "Calm Period (2004-2006)")
check_assumptions(x0[, tt_crisis], "Crisis Period (2007-2009)")
```

# LASSO-based FNETS for both Calm and Crisis Period (Sparsity test included) (n = 252)

```{r}

## FNETS sets

### Full FENTS forecast via LASSO

# (this little part is for the Sparsity Assumption test)
edge_density <- rep(NA, dim(x0)[2])
s_in_vec     <- rep(NA, dim(x0)[2])
q_vec        <- rep(NA, dim(x0)[2])
#===========

fl <- array(NA, dim = c(dim(x0), 1)) # 46*3269*1 array storing the one-step-ahead predicted values for each of the 46 firms on each trading day.
fle <- array(NA, dim = c(dim(fl)[2:3], 2)) # 3269*1*2 array storing the corresponding forecast errors, where the last dimension indexes two error metrics: avg (average relative squared error across all firms) and max (maximum relative absolute error across all firms)

dimnames(fl)[[3]] <-
  dimnames(fle)[[2]] <- 'lasso1.restricted'

dimnames(fle)[[3]] <- c('avg', 'max')

tt.ind <- c(tt_calm, tt_crisis)

## Main FNETS with only Factor Components and LASSO

for(tt in tt.ind){  # tt: forecast day index

  yl <- matrix(0, nrow = p, ncol = 1) # 46*1

  int <- max(1, tt - n):(tt - 1) # estimation window index, (tt-252):(tt-1), 252 trading days before the forecast day.  max(1, ...) preventing over the boundaries.
  
  x <- x0[, int]  # 46*252 matrix, window data
  x.true <- x0[, tt] # 46*1 array, the true value of tt day.
  
  # as FNETS assumes zero-mean process, so we need to demean first.
  mean.x <- apply(x, 1, mean) 
  xx <- x - mean.x
  
  ### FNETS Step 1
  
  dpca <- fnets:::dyn.pca(xx, q = NULL, ic.op = 5, kern.bw = kern.bw) # Estimates the spectral density matrix of the demeaned observations xx using the Barlett kernel estimator defined in Equation (7) in Barigozzi et al. (2024).  It Extracts the autocovariance matrices of X_t, chi_t and xi_t.  It's key output is 'acvGamma_i' which is the estimated autocovariance matrix of the latent idiosyncratic process xi_t, obtained by hat_Gamma_xi = hat_Gamma_x-hat_Gamma_chi.
  
  # setting q = NULL with ic.op = 5 instructs dyn.cpa() to estimate the number of common factors q from the data using the information criterion of Hallin & Liska (2007).
  
  
  q <- dpca$q
  
  if(is.na(q) | q == 0){  # Safeguard
    dpca <- fnets:::dyn.pca(xx, q = 1, ic.op = 5, kern.bw = kern.bw)
    q <- dpca$q
  }
  
  
  spec <- dpca$spec # estimated spectral density matrix which is the direct output of the Barlett kernel estimator.
  
  acv <- dpca$acv # estimated autocovaraince matrices of 'acvGamma_x' (sample autocovariance matrix of the observed process X_t), 'acvGamma_c' (estimated autocovariance matrix of the common component chi_t), 'acvGamma_i' (estimated autocovariance matrix of the latent idiosyncratic process xi_t)

  cve <- fnets:::common.irf.estimation(xx, Gamma_c = acv$Gamma_c, q = q,
                                       factor.var.order = NULL, max.var.order = NULL,
                                       trunc.lags = 20, n.perm = 30)  # Impulse response filter B(L) in the factor model form Equation (1) of the paper.  It fits a VAR to the estimated factors to obtain the dynamic factor loadings and truncated at 'trunc.lags=20' lags.
  
  obj <- list(acv = acv, spec = spec, loadings = cve$irf.est, factors = cve$u.est, q = q, mean.x = mean.x, kern.bw = kern.bw) 
  
  attr(obj, "factor") <- "unrestricted"
  
  # Here are all Step 1 outputs to be collected into a single list obj that is passed to the downstream functions.  attr(obj, "factor") <- "unrestricted" flags to common.preditc() that the common component was estimated under the unrestricted GDFM framework rather than a static factor model.
  
    cpre.static <- fnets:::common.predict(obj, x, n.ahead = 1, r = 'ic', fc.restricted = TRUE)
    # This produces the one-step-ahead forecast of the common component with obj.  fc.restricted = TRUE selects the restricted static facotr model forecasting method described in Section 4.1 of the original paper.
    
  #=================================================
  
  ## FNETS Step 2+3: Yule-Walker VAR + idiosyncratic forecast

  lambda.max <- max(abs(xx %*% t(xx)/n)) * 1
  lambda.path <- round(exp(seq(log(lambda.max), log(lambda.max * .0001), length.out = 20)), digits = 10)
  lambda <- lambda.path[9]
  
  dd <- 1
  mg <- fnets:::make.gg(acv$Gamma_i, dd)
  ive <- fnets:::var.lasso(mg$GG, mg$gg, lambda)
  
  # This part is the Section 3.3. of the original paper.  The VAR parameters beta are estimated via the l1-regularized Yule-Walker estimator in Equation (8).
  
  # Lambda is the LASSO penalization parameter in Equation (8).  It controls the sparsity of the estimated network.  The code uses a fixed grid of 20 values on a log scale between lambda.max and lambda.max * 0.0001 and picks the 9th value.
  
  # make.gg() constructs the Yule-Walker matrices G and g from the estimated autocovariance matrices of the latent idiosyncratic process obtained in Step 1.
  
  # var.lasso() solves the penalized problem and returns the estimated VAR transition matrix and related quantities.
  
  obj$idio.var <- ive
  ipre <- fnets:::idio.predict(obj, x, cpre.static, n.ahead = 1)
  yl[, 1] <- mean.x + cpre.static$fc + ipre$fc
  
  # Here is the forcasting.  Corresponding to Section 4 of the original paper.
  
  # idio.predict() computes the idiosyncratic forecast using the estimated VAR transition matrix.
  
  # yl is the final forecast adds back three components including sample mean, common component forecast and the idiosyncratic forecast.  
  
  # === Sparsity Assumption Test ===
  G <- ive$beta != 0
  edge_density[tt] <- sum(G) / (p * p)
  s_in_vec[tt]     <- max(colSums(G))
  q_vec[tt]        <- q
  # ===================================
  
  
  fl[, tt, ] <- yl
  fle[tt,, 1] <- apply(yl, 2, function(yy){sum((x.true - yy)^2)/sum(x.true^2)})
  fle[tt,, 2] <- apply(yl, 2, function(yy){max(abs(x.true - yy))/max(abs(x.true))})
  
  # Forecast errors
  # fl stores the raw forecast yl for day tt.
  # fle stores two relative forecast error metrics comparing the forecast yl against the true value x.ture.
  
}

ls <- list(fl = fl, fle = fle, edge_density = edge_density, s_in_vec = s_in_vec, q_vec = q_vec)
save(ls, file = paste('fnets_forecast.RData', sep = ''))


```


# Robustness (n = 126)


```{r}

# Robustness Check: n = 126 (half-year rolling window)

n126 <- 126
kern.bw126 <- floor(4 * (n126/log(n126))^(1/3))

fl126 <- array(NA, dim = c(dim(x0), 1))
fle126 <- array(NA, dim = c(dim(fl126)[2:3], 2))
dimnames(fl126)[[3]] <- dimnames(fle126)[[2]] <- 'lasso1.restricted'
dimnames(fle126)[[3]] <- c('avg', 'max')

# (this little part is for the Sparsity Assumption test)
edge_density126 <- rep(NA, dim(x0)[2])
s_in_vec126     <- rep(NA, dim(x0)[2])
q_vec126        <- rep(NA, dim(x0)[2])
#===========

for(tt in tt.ind){
  yl <- matrix(0, nrow = p, ncol = 1)
  int <- max(1, tt - n126):(tt - 1)
  x <- x0[, int]
  x.true <- x0[, tt]
  mean.x <- apply(x, 1, mean)
  xx <- x - mean.x

  dpca <- fnets:::dyn.pca(xx, q = NULL, ic.op = 5, kern.bw = kern.bw126)
  q <- dpca$q
  if(is.na(q) | q == 0){
    dpca <- fnets:::dyn.pca(xx, q = 1, ic.op = 5, kern.bw = kern.bw126)
    q <- dpca$q
  }
  spec <- dpca$spec
  acv  <- dpca$acv

  cve <- fnets:::common.irf.estimation(xx, Gamma_c = acv$Gamma_c, q = q,
                                       factor.var.order = NULL, max.var.order = NULL,
                                       trunc.lags = 20, n.perm = 30)

  obj <- list(acv = acv, spec = spec, loadings = cve$irf.est,
              factors = cve$u.est, q = q, mean.x = mean.x, kern.bw = kern.bw126)
  attr(obj, "factor") <- "unrestricted"

  cpre.static <- fnets:::common.predict(obj, x, n.ahead = 1, r = 'ic', fc.restricted = TRUE)

  lambda.max  <- max(abs(xx %*% t(xx)/n126)) * 1
  lambda.path <- round(exp(seq(log(lambda.max), log(lambda.max * .0001), length.out = 20)), digits = 10)
  lambda <- lambda.path[9]

  dd <- 1
  mg  <- fnets:::make.gg(acv$Gamma_i, dd)
  ive <- fnets:::var.lasso(mg$GG, mg$gg, lambda)
  obj$idio.var <- ive
  ipre <- fnets:::idio.predict(obj, x, cpre.static, n.ahead = 1)
  yl[, 1] <- mean.x + cpre.static$fc + ipre$fc
  
  
# === Sparsity Test ===
  G <- ive$beta != 0
  edge_density126[tt] <- sum(G) / (p * p)
  s_in_vec126[tt]     <- max(colSums(G))
  q_vec126[tt]        <- q
# ==========================

  fl126[, tt, ]  <- yl
  fle126[tt,, 1] <- apply(yl, 2, function(yy){ sum((x.true - yy)^2)/sum(x.true^2) })
  fle126[tt,, 2] <- apply(yl, 2, function(yy){ max(abs(x.true - yy))/max(abs(x.true)) })
}

ls126 <- list(fl = fl126, fle = fle126,
              edge_density = edge_density126,
              s_in_vec = s_in_vec126, q_vec = q_vec126)
save(ls126, file = 'fnets_forecast_126.RData')
```


# Factor number estimation

```{r}

# Factor number estimation (Assumption 1: Factor Strength)

x_calm   <- x0[, tt_calm]
x_crisis <- x0[, tt_crisis]

kern.bw_calm <- floor(4 * (ncol(x_calm)/log(ncol(x_calm)))^(1/3))
dpca_calm <- fnets:::dyn.pca(x_calm - rowMeans(x_calm), 
                              q = NULL, ic.op = 5, kern.bw = kern.bw_calm)

kern.bw_crisis <- floor(4 * (ncol(x_crisis)/log(ncol(x_crisis)))^(1/3))
dpca_crisis <- fnets:::dyn.pca(x_crisis - rowMeans(x_crisis), 
                                q = NULL, ic.op = 5, kern.bw = kern.bw_crisis)

cat("Calm period q:", dpca_calm$q, "\n")
cat("Crisis period q:", dpca_crisis$q, "\n")

```


# Network Structure (discarded, not use because time constrainted)


```{r}

# Network Structure: Full fnets() on calm and crisis periods

#library(igraph)

# Use the full period data (not rolling window)
#x_calm   <- x0[, tt_calm]
#x_crisis <- x0[, tt_crisis]

# Run full fnets for each period
#net_calm <- fnets(x_calm,
#                  q = "ic", var.order = 1, var.method = #"lasso",
#                  do.lrpc = TRUE,
#                  var.args = list(n.cores = 1))

#net_crisis <- fnets(x_crisis,
#                    q = "ic", var.order = 1, var.method = #"lasso",
#                    do.lrpc = TRUE,
#                    var.args = list(n.cores = 1))

# Extract adjacency matrices
# N^G: Granger (directed)
#G_calm   <- (net_calm$idio.var$beta   != 0) * 1
#G_crisis <- (net_crisis$idio.var$beta != 0) * 1

# N^C: Contemporaneous (undirected, from Gamma^{-1})
#C_calm   <- (net_calm$idio.var$Gamma_inv  != 0) * 1
#C_crisis <- (net_crisis$idio.var$Gamma_inv != 0) * 1
#diag(C_calm) <- diag(C_crisis) <- 0

# Network summary
#cat("=== Granger Network (N^G) ===\n")
#cat("Calm   - density:", round(sum(G_calm)   / (p*(p-1)), 4),
#    " | max in-degree:", max(colSums(G_calm)), "\n")
#cat("Crisis - density:", round(sum(G_crisis) / (p*(p-1)), 4),
#    " | max in-degree:", max(colSums(G_crisis)), "\n")

#cat("\n=== Contemporaneous Network (N^C) ===\n")
#cat("Calm   - density:", round(sum(C_calm)   / (p*(p-1)), 4),
#    " | max degree:", max(rowSums(C_calm)), "\n")
#cat("Crisis - density:", round(sum(C_crisis) / (p*(p-1)), 4),
#    " | max degree:", max(rowSums(C_crisis)), "\n")

# Top hub nodes (firms with highest in-degree in Granger network)
#firm_names <- colnames(old_clean)[-1]
#cat("\nTop 5 Granger hubs (calm):\n")
#print(sort(colSums(G_calm), decreasing=TRUE)[1:5])
#cat("Top 5 Granger hubs (crisis):\n")
#print(sort(colSums(G_crisis), decreasing=TRUE)[1:5])

```


# Results

### n = 252


```{r}

## read

load('fnets_forecast.RData')

for(period in c('calm', 'crisis')){
  tt.ind <- if(period == 'calm') tt_calm else tt_crisis
  err_avg <- ls$fle[tt.ind, 1, 1]  
  err_max <- ls$fle[tt.ind, 1, 2] 
  cat('\n', period, '\n')
  cat('avg - mean:', round(mean(err_avg, na.rm=TRUE), 4),
      'median:', round(median(err_avg, na.rm=TRUE), 4),
      'sd:', round(sd(err_avg, na.rm=TRUE), 4), '\n')
  cat('max - mean:', round(mean(err_max, na.rm=TRUE), 4),
      'median:', round(median(err_max, na.rm=TRUE), 4),
      'sd:', round(sd(err_max, na.rm=TRUE), 4), '\n')
}

cat('=== Calm: sparsity ===\n')
cat('Edge density mean:', round(mean(ls$edge_density[tt_calm], na.rm=TRUE), 4), '\n')
cat('s_in mean:',        round(mean(ls$s_in_vec[tt_calm],     na.rm=TRUE), 2), '\n')
cat('q mean:',           round(mean(ls$q_vec[tt_calm],         na.rm=TRUE), 2), '\n')

cat('=== Crisis: sparsity ===\n')
cat('Edge density mean:', round(mean(ls$edge_density[tt_crisis], na.rm=TRUE), 4), '\n')
cat('s_in mean:',        round(mean(ls$s_in_vec[tt_crisis],     na.rm=TRUE), 2), '\n')
cat('q mean:',           round(mean(ls$q_vec[tt_crisis],         na.rm=TRUE), 2), '\n')

```


### n = 126

```{r}

# Robustness Check: n=126 results
load('fnets_forecast_126.RData')

for(period in c('calm', 'crisis')){
  tt.ind <- if(period == 'calm') tt_calm else tt_crisis
  err_avg <- ls126$fle[tt.ind, 1, 1]
  err_max <- ls126$fle[tt.ind, 1, 2]
  cat('\n[n=126]', period, '\n')
  cat('avg - mean:', round(mean(err_avg, na.rm=TRUE), 4),
      'median:', round(median(err_avg, na.rm=TRUE), 4),
      'sd:', round(sd(err_avg, na.rm=TRUE), 4), '\n')
  cat('max - mean:', round(mean(err_max, na.rm=TRUE), 4),
      'median:', round(median(err_max, na.rm=TRUE), 4),
      'sd:', round(sd(err_max, na.rm=TRUE), 4), '\n')
}

cat('=== [n=126] Calm: sparsity ===\n')
cat('Edge density mean:', round(mean(ls126$edge_density[tt_calm], na.rm=TRUE), 4), '\n')
cat('s_in mean:',        round(mean(ls126$s_in_vec[tt_calm],     na.rm=TRUE), 2), '\n')
cat('q mean:',           round(mean(ls126$q_vec[tt_calm],         na.rm=TRUE), 2), '\n')

cat('=== [n=126] Crisis: sparsity ===\n')
cat('Edge density mean:', round(mean(ls126$edge_density[tt_crisis], na.rm=TRUE), 4), '\n')
cat('s_in mean:',        round(mean(ls126$s_in_vec[tt_crisis],     na.rm=TRUE), 2), '\n')
cat('q mean:',           round(mean(ls126$q_vec[tt_crisis],         na.rm=TRUE), 2), '\n')

```


# Result Significance

```{r}

## t-test on the FNETS forcast results (n=252)

fe_avg_calm <- ls$fle[tt_calm, 1, 1]
fe_avg_crisis <- ls$fle[tt_crisis, 1, 1]

fe_max_calm <- ls$fle[tt_calm, 1, 2]
fe_max_crisis <- ls$fle[tt_crisis, 1, 2]

t.test(fe_avg_calm, fe_avg_crisis, alternative = "two.sided")
t.test(fe_max_calm, fe_max_crisis, alternative = "two.sided")

# n=126 t-tests
fe_avg_calm126   <- ls126$fle[tt_calm,   1, 1]
fe_avg_crisis126 <- ls126$fle[tt_crisis, 1, 1]

fe_max_calm126   <- ls126$fle[tt_calm,   1, 2]
fe_max_crisis126 <- ls126$fle[tt_crisis, 1, 2]

t.test(fe_avg_calm126, fe_avg_crisis126, alternative = "two.sided")
t.test(fe_max_calm126, fe_max_crisis126, alternative = "two.sided")

```

# Plots

```{r}
# Plotting


library(ggplot2)
library(patchwork)

dates_calm   <- dates[tt_calm]
dates_crisis <- dates[tt_crisis]

all_tt    <- c(tt_calm, tt_crisis)
all_dates <- dates[all_tt]
period_label <- c(rep("Calm (2004-2006)", length(tt_calm)),
                  rep("Crisis (2007-2009)", length(tt_crisis)))

# =============
# FE_max time series plot


df_femax <- data.frame(
  date   = all_dates,
  fe_max = ls$fle[all_tt, 1, 2],
  period = period_label
)

p1 <- ggplot(df_femax, aes(x = date, y = fe_max, color = period)) +
  geom_line(linewidth = 0.4, alpha = 0.8) +
  geom_vline(xintercept = as.Date("2007-01-01"), 
             linetype = "dashed", color = "grey40", linewidth = 0.6) +
  scale_color_manual(values = c("Calm (2004-2006)"   = "#2166ac",
                                "Crisis (2007-2009)" = "#d6604d")) +
  labs(title = "Figure 2: Max Forecast Error Over Time",
       x = NULL, y = "FE_max", color = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 11))

# ==============
# Edge Density Time series plot

df_edge <- data.frame(
  date         = all_dates,
  edge_density = ls$edge_density[all_tt],
  period       = period_label
)

p2 <- ggplot(df_edge, aes(x = date, y = edge_density, color = period)) +
  geom_line(linewidth = 0.4, alpha = 0.8) +
  geom_vline(xintercept = as.Date("2007-01-01"),
             linetype = "dashed", color = "grey40", linewidth = 0.6) +
  scale_color_manual(values = c("Calm (2004-2006)"   = "#2166ac",
                                "Crisis (2007-2009)" = "#d6604d")) +
  labs(title = "Figure 1: Network Edge Density Over Time",
       x = NULL, y = "Edge Density", color = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 11))

# ============
# FE_avg Box plot

df_avg <- data.frame(
  value  = c(ls$fle[tt_calm,     1, 1], ls$fle[tt_crisis,     1, 1],
             ls126$fle[tt_calm,  1, 1], ls126$fle[tt_crisis,  1, 1]),
  period = c(rep("Calm",   length(tt_calm)),   rep("Crisis", length(tt_crisis)),
             rep("Calm",   length(tt_calm)),   rep("Crisis", length(tt_crisis))),
  window = c(rep("n = 252", length(tt_calm) + length(tt_crisis)),
             rep("n = 126", length(tt_calm) + length(tt_crisis)))
)

p3a <- ggplot(df_avg, aes(x = period, y = value, fill = period)) +
  geom_boxplot(outlier.size = 0.3, outlier.alpha = 0.3) +
  facet_wrap(~ window) +
  scale_fill_manual(values = c("Calm" = "#92c5de", "Crisis" = "#f4a582")) +
  labs(title = "Figure 4: Average Forecast Error by Period and Window Size",
       x = NULL, y = "FE_avg", fill = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 11))

# ==============
# FE_max Box plot

df_max <- data.frame(
  value  = c(ls$fle[tt_calm,     1, 2], ls$fle[tt_crisis,     1, 2],
             ls126$fle[tt_calm,  1, 2], ls126$fle[tt_crisis,  1, 2]),
  period = c(rep("Calm",   length(tt_calm)),   rep("Crisis", length(tt_crisis)),
             rep("Calm",   length(tt_calm)),   rep("Crisis", length(tt_crisis))),
  window = c(rep("n = 252", length(tt_calm) + length(tt_crisis)),
             rep("n = 126", length(tt_calm) + length(tt_crisis)))
)

p3b <- ggplot(df_max, aes(x = period, y = value, fill = period)) +
  geom_boxplot(outlier.size = 0.3, outlier.alpha = 0.3) +
  facet_wrap(~ window) +
  scale_fill_manual(values = c("Calm" = "#92c5de", "Crisis" = "#f4a582")) +
  labs(title = "Figure 5: Max Forecast Error by Period and Window Size",
       x = NULL, y = "FE_max", fill = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 11))

# ============
# Outputs

ggsave("fig1_femax_timeseries.pdf", p1, width = 8, height = 3.5)
ggsave("fig2_edge_density.pdf",     p2, width = 8, height = 3.5)
ggsave("fig3_boxplot_avg.pdf", p3a, width = 6, height = 4)
ggsave("fig4_boxplot_max.pdf", p3b, width = 6, height = 4)

```

```{r}
# ============
# Cross-sectional mean log-volatility time series (Stationary figure)

mean_calm_ts   <- colMeans(x0[, tt_calm])
mean_crisis_ts <- colMeans(x0[, tt_crisis])

df_mean <- data.frame(
  date   = c(dates[tt_calm], dates[tt_crisis]),
  value  = c(mean_calm_ts, mean_crisis_ts),
  period = c(rep("Calm (2004-2006)",   length(tt_calm)),
             rep("Crisis (2007-2009)", length(tt_crisis)))
)

p_mean <- ggplot(df_mean, aes(x = date, y = value, color = period)) +
  geom_line(linewidth = 0.4, alpha = 0.8) +
  geom_hline(yintercept = mean(mean_calm_ts),   linetype = "dashed",
             color = "#2166ac", linewidth = 0.5) +
  geom_hline(yintercept = mean(mean_crisis_ts), linetype = "dashed",
             color = "#d6604d", linewidth = 0.5) +
  scale_color_manual(values = c("Calm (2004-2006)"   = "#2166ac",
                                "Crisis (2007-2009)" = "#d6604d")) +
  labs(title = "Figure 3: Cross-Sectional Mean Log-Volatility",
       subtitle = "Dashed lines indicate period means",
       x = NULL, y = "Mean log-volatility", color = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 11))

ggsave("fig0_mean_logvol.pdf", p_mean, width = 8, height = 3.5)
```

# Reference

1. Barigozzi M, Cho H, Owens D (2024). fnets: Factor-Adjusted Network Estimation and Forecasting for High-Dimensional Time Series. doi:10.32614/CRAN.package.fnets <https://doi.org/10.32614/CRAN.package.fnets>, R package version
  0.1.6, <https://CRAN.R-project.org/package=fnets>.

2. Komsta L, Novomestky F (2022). moments: Moments, Cumulants, Skewness, Kurtosis and Related Tests.
  doi:10.32614/CRAN.package.moments <https://doi.org/10.32614/CRAN.package.moments>, 
  R package version 0.14.1, <https://CRAN.R-project.org/package=moments>.

3. Trapletti A, Hornik K (2026). tseries: Time Series Analysis and Computational Finance.
  doi:10.32614/CRAN.package.tseries <https://doi.org/10.32614/CRAN.package.tseries>, R package version 0.10-61,
  <https://CRAN.R-project.org/package=tseries>.